### 2017-06-20 sec-conclusion.tex
Minor revision to related work, the parameter estimation section, conclusions, and acknowledgments.

\section{Conclusion}
% summarize the main contributions of the paper; summarize the experimental
% results; argue again for our novelty; discuss future directions
As a tool to help instructors and eduation researchers better understand behavior of MOOC students, 
 we proposed a two-layer hidden Markov model to automatically extract 
student activity patterns in the form of behavior state-transition graphs from large amounts of 
MOOC log data. This model is different from existing
methods in that it treats behavior patterns as a sequence of \emph{latent
states}, rather than assigning these states in a rule-based manner. It
captures the variable behaviors of students over time, and allows
analysis at different levels of granularity.

@@

\subsection{Future Work}

Although we only experimented with our model on two MOOCs, the model is
completely general and can be easily applied to analyze the log of any
other course to enable deep understanding of student behaviors as well as
the correlations of such behaviors and other variables such as grades. We
plan to develop a MOOC log analysis system based on the proposed model to
both facilitate education research and help instructors improve course
design.

Our model can empower many new comparative analyses. For example, we could
now look at how behavior patterns change between different offerings of
the same MOOC to understand how changes in course structure or materials
influences student behavior. Individual students can now also be compared
against each other or against groups. For example, by decoding the latent
state sequences for each student, we can measure how ``surprising'' their
latent state transition sequence is relative to the average we would expect
according to the model, or to the average ``perfect'' student, etc. We can
now investigate how certain behavioral patterns correlate with properties
of a student (e.g., demographics, prior aptitude, etc.). After decoding the
students' latent state sequences, we could also correlate course-wide
drifts in these latent states with events in a course. For example, we
might be able to automatically discover difficult or confusing parts of a
course by noticing spikes in the distribution of students over latent
states over time.

Our model does not explicitly model drop-out like
\citet{Kizilcec:2013:LAK}, but doing so is an obvious extension. Our model
should be able to provide predictions of when a student is ``at risk'' for
dropping out under such an extension.

Currently, the model learns a transition matrix over the latent states that
is \emph{shared} across all students. It would be interesting to instead
learn a different latent state transition matrix for each individual
student, but keep the second-level Markov models shared. This would provide
the model with more flexibility which may be desirable itself, but would
also naturally result in a description of a student (via his or her HMM
transitions) that could be incorporated into existing supervised learning
techniques that try to predict student outcomes for understanding
which of the latent behavior patterns discovered by 2L-HMM are most predictive
of the performance of student learning. One could also relax this somewhat
and posit that \emph{groups} of students transition between the lower layer
patterns our method identifies in distinct ways and instead do a soft
clustering of students into $K_2$ groups based on the similarity of their
transitioning behavior between the higher-level behaviors we can identify.

There is more recent work on better learning algorithms for mixtures of
Markov models~\citep{Gupta:2016:NIPS}. It would be worth exploring whether
the advances proposed in this and similar work can be applied to our model
to address some of the concerns surrounding our use of the EM algorithm for
our parameter estimation.

>>>>

\section{Conclusions and Future Work}
% summarize the main contributions of the paper; summarize the experimental
% results; argue again for our novelty; discuss future directions
As a tool to help instructors and eduation researchers better understand behavior of MOOC students, 
 we proposed a two-layer hidden Markov model to automatically extract 
student activity patterns in the form of behavior state-transition graphs from large amounts of 
MOOC log data. This model is different from existing
methods in that it treats behavior patterns as a sequence of \emph{latent
states}, rather than assigning these states in a rule-based manner. It
captures the variable behaviors of students over time, and allows
analysis at different levels of granularity.

@@

%\subsection{Future Work}

There are a number of interesting directions to further extend our work in the future. 
First, the proposed model is 
completely general and can thus be easily applied to analyze the log of many
other courses to enable deep understanding of student behaviors as well as
the correlations of such behaviors and other variables such as grades. To realize these benefits, 
it would be useful 
to develop a MOOC log analysis system based on the proposed model so as to facilitate education research and help instructors improve course
design.

Second, our model can empower many new comparative analyses. For example, we could
now look at how behavior patterns change between different offerings of
the same MOOC to understand how changes in course structure or materials
influences student behavior. Individual students can now also be compared
against each other or against groups. For example, by decoding the latent
state sequences for each student, we can measure how ``surprising'' their
latent state transition sequence is relative to the average we would expect
according to the model, or to the average ``perfect'' student, etc. We can
now investigate how certain behavioral patterns correlate with properties
of a student (e.g., demographics, prior aptitude, etc.). After decoding the
students' latent state sequences, we could also correlate course-wide
drifts in these latent states with events in a course. For example, we
might be able to automatically discover difficult or confusing parts of a
course by noticing spikes in the distribution of students over latent
states over time.

Third, there is more recent work on better learning algorithms for mixtures of
Markov models~\citep{Gupta:2016:NIPS}. It would be worth exploring whether
the advances proposed in this and similar work can be applied to our model
to address some of the concerns surrounding our use of the EM algorithm for
our parameter estimation.

Finally, the proposed model can be extended in several ways. For example, 
although our model does not explicitly model drop-out like
\citet{Kizilcec:2013:LAK}, doing so is an obvious extension. Our model
would be able to provide predictions of when a student is ``at risk'' for
dropping out under such an extension.
Also, currently, the model learns a transition matrix over the latent states that
is \emph{shared} across all students. It would be interesting to instead
learn a different latent state transition matrix for each individual
student, but keep the second-level Markov models shared. This would provide
the model with more flexibility which may be desirable itself, but would
also naturally result in a description of a student (via his or her HMM
transitions) that could be incorporated into existing supervised learning
techniques that try to predict student outcomes for understanding
which of the latent behavior patterns discovered by 2L-HMM are most predictive
of the performance of student learning. One could also relax this somewhat
and posit that \emph{groups} of students transition between the lower layer
patterns our method identifies in distinct ways and instead do a soft
clustering of students into $K_2$ groups based on the similarity of their
transitioning behavior between the higher-level behaviors we can identify.

### 2017-06-20 sec-model.tex
Minor revision to related work, the parameter estimation section, conclusions, and acknowledgments.

\subsection{Parameter Estimation}
\label{sec:param-estimation}
To learn the parameters of our model, we may attempt to use maximum
likelihood estimation. Unfortunately, a closed-form solution does not
exist, so we must appeal to the EM algorithm~\cite{Dempster:1977:JRSS}. In
particular, we propose a minor modification of the Baum-Welch
algorithm~\cite{Rabiner:1990:RSR} which is used to learn the parameters for
hidden Markov models in an unsupervised setting. In the following sections,
we will provide a brief description of the original Baum-Welch algorithm
for unsupervised parameter estimation for categorical valued hidden Markov
models, and then describe our modification to allow for parameter
estimation for our 2L-HMM modification for sequence valued observations.

>>>>

\subsection{Parameter Estimation}
\label{sec:param-estimation}
To learn the parameters of our model, we may attempt to use maximum
likelihood estimation. Unfortunately, a closed-form solution does not
exist, so we must appeal to the EM algorithm~\cite{Dempster:1977:JRSS}. In
particular, we propose a minor modification of the Baum-Welch
algorithm~\cite{Rabiner:1990:RSR} which is an efficient EM algorithm
for learning the parameters for
hidden Markov models in an unsupervised setting where the Markovian assumption is exploited
 to significantly reduce the computational complexity of the EM algorithm by avoiding explicit enumeration of all the possible state transitions.  In the following sections,
we will provide a brief description of the original Baum-Welch algorithm
for unsupervised parameter estimation for categorical valued hidden Markov
models, and then describe our modification to allow for parameter
estimation for our 2L-HMM modification for sequence valued observations.

### 2017-06-20 sec-related-work.tex
Minor revision to related work, the parameter estimation section, conclusions, and acknowledgments.

Our formulation differs from both in that we do not feed the labeled
sequence of the lower level into the input of the higher level. Instead,
our lower level is actually treated using a non-hidden Markov model, and
the higher level is modeling transitions between the $K$ different
non-hidden Markov models we consider. The problem to be solved is similar
in that we wish to predict a ``label'' for a sequence of actions a student
takes as well as understand the transition behavior between those labels.
However, one of the consequences of modeling the lower layer using a
non-hidden Markov model instead of an HMM directly is that the meaning of
the $K$ different latent states can be more clearly captured by using our
proposed behavior representation. If we were to use an HMM as our first
layer, the behavior patterns (like from
Figure~\ref{fig:motivating-example}) would instead have nodes that
represent another set of latent states instead of being concrete actions
themselves. Each of these latent states would then be associated with some
other output distribution over the possible concrete actions to be
considered. In order to understand a single behavior pattern uncovered, one
would first have to understand the different output distributions for the
latent states in that pattern in order to understand the meaning of that
latent state. This further complicates the understanding of higher level
patterns because understanding the higher layer patterns requires
understanding the lower layer patterns. By taking a more restrictive view
on the first layer, we are able to produce a representation that can be
more readily interpreted due to the states in our first layer
representation having an immediately clear meaning (the concrete action
they represent).

>>>>

Our formulation differs from the work~\cite{Zhang:2004:CVPR,Oliver:2004:CVIU} in that we do not feed the labeled
sequence of the lower level into the input of the higher level. Instead,
our lower level is actually treated using a non-hidden Markov model, and
the higher level is modeling transitions between the $K$ different
non-hidden Markov models we consider. The problem to be solved is similar
in that we wish to predict a ``label'' for a sequence of actions a student
takes as well as understand the transition behavior between those labels.
However, one of the consequences of modeling the lower layer using a
non-hidden Markov model instead of an HMM directly is that the meaning of
the $K$ different latent states can be more clearly captured by using our
proposed behavior representation. If we were to use an HMM as our first
layer, the behavior patterns (like from
Figure~\ref{fig:motivating-example}) would instead have nodes that
represent another set of latent states instead of being concrete actions
themselves. Each of these latent states would then be associated with some
other output distribution over the possible concrete actions to be
considered. In order to understand a single behavior pattern uncovered, one
would first have to understand the different output distributions for the
latent states in that pattern in order to understand the meaning of that
latent state. This further complicates the understanding of higher level
patterns because understanding the higher layer patterns requires
understanding the lower layer patterns. By taking a more restrictive view
on the first layer, we are able to produce a representation that can be
more readily interpreted due to the states in our first layer
representation having an immediately clear meaning (the concrete action
they represent).
